ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-04 rebuttal.tex

\begin{reviewer}
    %
    24.~The Python code snippets are not always initiated in the text.
    %
\end{reviewer}

\begin{response}
    %
    We have corrected this oversight.
    %
\end{response}

\begin{reviewer}
    %
    25.~Page 10, line 55: add more commentary text. It is not clear that those lines create the
    necessary instances for Cantera, rather than ChemKED.
    %
\end{reviewer}

\begin{response}
    %
    We have expanded the explanation for all of the Python examples to give more detail of the
    process of creating the simulation.
    %
\end{response}

@@

\begin{response}
    %
    \verb|air.xml| is an input file included with Cantera, representing an average composition of
    air. We have modified the text to include this information.
    %
\end{response}

\subsubsection*{Shock tube modelling with constant volume}

\begin{reviewer}
    %
    27.~The table overlapping page 11 and 12 can be reduced by truncating the \texttt{DataPoints},
    similarly as is done with the table starting at page 9, line 31 and ending at page 10, line 27.
    %
\end{reviewer}

\begin{response}
    %
    We have truncated the datapoints list as the reviewer suggests.
    %
\end{response}

@@

\begin{reviewer}
    %
    28. Python code snippet on page 13: (a.) Add more commentary text. (b.) Changing the example to
    enable multithreading might make it more appealing to start using ChemKED as a data format in
    addition with Cantera as a simulation tool.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the example to include more explanatory text and to use Python's built-in
    multiprocessing library to run the simulations on many processors. We note that if Cantera has
    been built with multi-threaded BLAS and LAPACK libraries (such as the Intel MKL), then the 0-D
    solver will automatically utilize the multithreading capabilities of those libraries. Therefore,
    using the multiprocessing library may not give much benefit in terms of overall speedup,
    depending on the complexity of the chemical kinetic mechanism (i.e., how long the simulation
    takes to run), how many processors are available, and the availability of the multi-threaded
    BLAS and LAPACk libraries. However, we have not explained this detail of Cantera in the text to
    avoid confusing the reader.
    %
\end{response}

### 2017-10-04 pyked.tex
More work on example explanation
Also add multiprocessing. I haven't checked the code yet though! Work on
#44, #43, #42, #41, #40, #39

At this point, the user could post-process the information in the \pybox|pressure| list to calculate
the derivative by whatever algorithm they choose.

@@

The \ck{} file used in this example can be found in the \texttt{tests} directory of the \pk{}
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
%
This \ck{} file specifies multiple data points with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox|common-properties| section is specified.
%
\begin{yamlbox}
common-properties:
  composition: &comp
    kind: mole fraction
    species:
      - species-name: t-butanol
        InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
        amount:
          - 0.003333333
      - species-name: O2
        InChI:  1S/O2/c1-2
        amount:
          - 0.04
      - species-name: Ar
        InChI:  1S/Ar
        amount:
          - 0.956666667
  ignition-type: &ign
    target: OH*
    type: 1/2 max
datapoints:
  - temperature:
      - 1459 kelvin
    ignition-delay:
      - 347 us
    pressure:
      - 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1389 kelvin
    ignition-delay:
      - 756 us
    pressure:
      - 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1497 kelvin
    ignition-delay:
      - 212 us
    pressure:
      - 1.55 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1562 kelvin
    ignition-delay:
      - 105 us
    pressure:
      - 1.50 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
\end{yamlbox}

In this example, we would like to run constant-volume simulations at each
pressure and temperature condition in the \yabox|datapoints| list. Once again,
the \ck{} file specifies all the information required for the simulations except
for the chemical kinetic model, and Cantera can be used to simulate autoignition:
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

ck = ChemKED('Stranic2012-tbuoh.yaml')

gas = ct.Solution('tbuoh-mechanism.cti')

# Since the composition is specified by the common-properties,
# just take the composition of the first DataPoint
X_initial = ck.datapoints[0].get_cantera_mole_fraction()

# Loop through each of the DataPoints
for dp in ck.datapoints:
    T_initial = dp.temperature.to('K').magnitude
    P_initial = dp.pressure.to('Pa').magnitude
    gas.TPX = T_initial, P_initial, X_initial
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    # Define ignition delay as T_initial + 400 K
    while reac.T < T_initial + 400:
        netw.step()

    print('The ignition delay for T_initial={}, P_initial={} is: '
          '{} seconds'.format(T_initial, P_initial, netw.time)
          )
\end{pythonbox}
%
The ignition delay in this example is defined as equal to the \SI{400}{\kelvin}
above the initial temperature, a simplified definition used for this example.
In general, the user could post-process the concentration information from the
simulation to determine one-half the maximum \ce{OH^*} concentration to match
the experimental definition of ignition delay.
%====================================================================
\section{Conclusions and future work}

>>>>

At this point, the user would post-process the information in the \pybox|pressure| list to calculate
the derivative by whatever algorithm they choose.

@@

The \ck{} file used in this example can be found in the \texttt{tests} directory of the \pk{}
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in
Section~\ref{sec:rcm-modeling}.

This \ck{} file specifies multiple data points with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox|common-properties| section is specified, followed by the
datapoints list:
%
\begin{yamlbox}
common-properties:
  composition: &comp
    kind: mole fraction
    species:
      - species-name: t-butanol
        InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
        amount:
          - 0.003333333
      - species-name: O2
        InChI:  1S/O2/c1-2
        amount:
          - 0.04
      - species-name: Ar
        InChI:  1S/Ar
        amount:
          - 0.956666667
  ignition-type: &ign
    target: OH*
    type: 1/2 max
datapoints:
  - temperature:
      - 1459 kelvin
    ignition-delay:
      - 347 us
    pressure:
      - 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1389 kelvin
    ignition-delay:
      - 756 us
    pressure:
      - 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
    ...
\end{yamlbox}
%
As before, we have truncated the datapoints list for space.

In this example, we will run constant-volume simulations at each
pressure and temperature condition in the \yabox|datapoints| list. Once again,
the \ck{} file specifies all the information required for the simulations except
for the chemical kinetic model, and Cantera can be used to simulate autoignition. The setup steps
are the same as in the previous example, with one exception. In this example, we also import
Python's built-in multiprocessing library so that we can run the simulations on multiple cores. We
import the \pybox|Pool| class, which offers facilities to manage a jobs on a pool of processors:
%
\begin{pythonbox}
import cantera as ct
from multiprocessing import Pool
from pyked import ChemKED
\end{pythonbox}

Then, we define a function that will be mapped onto each job. This function takes the initial
temperature, pressure, and mole fractions as input and returns the simulated ignition delay time,
defined as the time the temperature increases by \SI{400}{\K} over the initial temperature, a
simplified definition for this example. In general, the user could process the mole fraction of
\ce{OH^*} (provided that the mechanism included accurate chemistry for \ce{OH^*}) to match the
definition of ignition delay in the experiments. As before, the \verb|tbuoh-mechanism.cti| file must
be supplied by the user:
%
\begin{pythonbox}
def run_simulation(T, P, X):
    gas = ct.Solution('tbuoh-mechanism.cti')
    gas.TPX = T, P, X
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    while reac.T < T + 400:
        netw.step()

    return netw.time
\end{pythonbox}

Then, we load the ChemKED file and generate a list of initial conditions that will be mapped onto
the \pybox|run_simulation()| function. We first define a convenience function to collect the input
from a single datapoint and return a Python tuple with the conditions (tuples are lightweight groups
of data in Python). Then, we use the built-in \pybox|map| function to apply the
\pybox|collect_input| function to each of the elements in the \pybox|ck.datapoints| list: convert a
\pybox|Quantity| to some input units and return the magnitude, then use Python's list comprehension
syntax to generate a list of tuples (tuples are lightweight groups of data in Python):
%
\begin{pythonbox}
ck = ChemKED('Stranic2012-tbuoh.yaml')

def collect_input(dp):
    T_initial = dp.temperature.to('K').magnitude
    P_initial = dp.pressure.to('Pa').magnitude
    X_initial = dp.get_cantera_mole_fraction()
    return (T_initial, P_initial, X_initial)

initial_conditions = map(collect_input, ck.datapoints)
initial_conditions = [(to_SI(dp.temperature, 'K'), to_SI(dp.pressure, 'Pa'), dp.get_cantera_mole_fraction()) for dp in ck.datapoints]
\end{pythonbox}

Finally, we create the processor \pybox|Pool| and send the jobs out to run:
%
\begin{pythonbox}
with Pool(12) as pool:
    ignition_delays = pool.starmap(run_simulation, initial_conditions)

for (T, P, X), tau in zip(initial_conditions, ignition_delays):
    print(f'The ignition delay for T_initial={T}, P_initial={P} is: {tau} seconds')
\end{pythonbox}
%
The simulated ignition delay results are returned in the \pybox|ignition_delays| list. The results
are printed to the screen using a Python formatted string (f-string).
%====================================================================
\section{Conclusions and future work}
